---
title: "In Class Exercise 4"
author: "edwin"
date: "September 09,2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

#### Loading R packages

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse,sparr)
```

#### Importing and preparing study area

```{r}
kbb <- st_read(dsn="data/rawdata",
               layer ="Kepulauan_Bangka_Belitung")
 
```

Revise code

```{r}
kbb_sf <- st_read(dsn="data/rawdata",
               layer ="Kepulauan_Bangka_Belitung") %>% 
  st_union() %>% #to dissovle
  st_zm(drop=TRUE,what="ZM") %>% #to drop the z from polygon
  st_transform(crs=32748)
```

#### Converting OWIN

```{r}
kbb_owin<-as.owin(kbb_sf)
kbb_owin
```

class() to confirm output is owin

```{r}
class(kbb_owin)
```

#### Import forest fire data

```{r}
fire_sf<-read.csv("data/rawdata/forestfires.csv") %>% 
  st_as_sf(coords = c("longitude","latitude"),
           crs=4326) %>% 
  st_transform(crs=32748)
```

because **ppp** object only accept numerical or character as mark, the code below is used to convert acq_date to numeric

```{r}
fire_sf<-fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>% 
  mutate(Month_num = month(acq_date)) %>% 
  mutate(Month_fac = month(acq_date,
                           label = TRUE,
                           abbr = FALSE))
```

#### Visualising

```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots()
```

##### Visualing by month

```{r}
tm_shape(kbb_sf)+
  tm_polygons()+
  tm_shape(fire_sf)+
  tm_dots(size=0.1)+
  tm_facets(by="Month_fac",
            free.coords = FALSE,#if true, zoom to where the data points are
            drop.units = TRUE)
```

#### Computing STKDE by Month

The code below remove the unwanted fields form fire_sf. This is because **as.ppp()** on need the mark field and geometry field from the input sf.dataframe

```{r}
fire_month<-fire_sf %>% 
  select(Month_num)
```

The code below is used to derive **ppp** object call fire_month from fire_month sf dataframe

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

the code below combine **origin_am_ppp** and **am_owin** objects into one

```{r}
fire_month_owin <-fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

#### Computing STKDE

**spattemp_density()** of sparr is used to compute STKDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
```

```{r}
summary(st_kde)
```

##### Plotting STKDE object

in the code below, plot() of R base is used to the KDE between july 2023-december 2023

```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))#2rows,3 col
for (i in tims){
  plot(st_kde,i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month",i))
} 
```
