---
title: "In Class Exercise 10"
author: "edwin"
---

```{r}
pacman::p_load(olsrr,ggstatsplot, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary,performance)
```

### **Importing the aspatial data**

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

### **Importing geospatial data**

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

### Correlation Analysis -ggstatsplot methods

```{r}
ggcorrmat(condo_resale[,5:23])
```

### Building a Hedonic Pricing Model by using Multiple Linear Regression Method

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

### Generating tidy linear regression report

```{r}
olsrr_condo<-ols_regress(condo.mlr)
```

### Variable selection

```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo.mlr,
  p_val = 0.05,
  details = FALSE
)
condo_fw_mlr
plot(condo_fw_mlr)
```

```{r}
condo_bw_mlr <- ols_step_backward_p(
  condo.mlr,
  p_val = 0.05,
  details = FALSE
)
condo_bw_mlr
plot(condo_bw_mlr)
```

```{r}
condo_bi_mlr <- ols_step_both_p(
  condo.mlr,
  p_val = 0.05,
  details = FALSE
)
condo_bi_mlr
plot(condo_bi_mlr)
```

### Model Selection

```{r}
metric <- compare_performance(condo.mlr,
                              condo_fw_mlr$model,
                              condo_bw_mlr$model,
                              condo_bi_mlr$model)
```

```{r}
metric$Name <- gsub(".*\\\\([a-zA-Z0-9_]+)\\\\, \\\\model\\\\.*","\\1",metric$Name)
plot(metric)
```

### Visualisng Model Paramenters

```{r}
ggcoefstats(condo_bi_mlr$model,
           sort="ascending")
```

### Check for multicollinearity

```{r}
check_collinearity(condo_bi_mlr$model)

plot(check_collinearity(condo_bi_mlr$model))+
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1
  ))
```

### Linearity assumption test

```{r}
out <-plot(check_model(condo_bi_mlr$model,
                       panel=FALSE))

out[[2]]
```

```{r}
plot(check_normality(condo_bi_mlr$model))
```

### Checking of outliers

```{r}
outliers <- check_outliers(condo_bi_mlr$model,
                           method='cook')

outliers
```

```{r}
plot(check_outliers(condo_bi_mlr$model,
                    method='cook'))
```

### Visualisng spatial non-stationary

```{r}
mlr_output<- as.data.frame(condo_fw_mlr$model$residuals) %>% 
  rename('FW_MLR_RES'='condo_fw_mlr$model$residuals')
```
