---
title: "Take Home Exercise 1: Geospatial Analytics for Social Good: Application of Spatial and Spatio-temporal Point Patterns Analysis to discover the geographical distribution of Armed Conflict in Myanmar"
author: "edwin tan"
execute: 
  eval: true
  echo: true
  freeze: true
---

# The Armed Conflict in Myanmar: A Complex and Ongoing Crisis

**Background**

The armed conflict in Myanmar, also known as the Burmese Civil War, is a decades-long struggle that has been exacerbated by recent political developments. The conflict primarily involves the Myanmar military (Tatmadaw) and various ethnic armed organizations (EAOs) representing different ethnic minorities

**Key Factors Contributing to the Conflict**

-   **Ethnic Tensions:** Deep-rooted ethnic tensions and historical grievances have been a major driving force behind the conflict. Many ethnic minorities, such as the Rohingya, Karen, and Shan, have long sought greater autonomy or independence from the central government.  

-   **Military Dominance:** The Tatmadaw has exercised significant control over Myanmar's political and military affairs for decades. Its authoritarian rule and suppression of dissent have fueled resentment among many ethnic groups.  

-   **Resource Competition:** Disputes over natural resources, such as timber, jade, and gas, have also contributed to the conflict. Control of these resources can provide significant economic and political advantages.

-   **External Influences:** Neighboring countries, such as China and Thailand, have sometimes been involved in the conflict, either directly or indirectly, through their support for different factions.

**Recent Developments**

The conflict escalated significantly following the 2021 military coup, which ousted the democratically elected government led by Aung San Suu Kyi.

The coup sparked widespread protests and civil disobedience, leading to a crackdown by the military. This crackdown has resulted in numerous casualties and human rights abuses.  

In response to the coup, a number of armed resistance groups have emerged, including the People's Defense Forces (PDFs), which are civilian-led militias fighting against the military. These groups have engaged in clashes with the Tatmadaw across the country.

# Loading Library

-   **sf:** This package provides simple features for representing spatial data in R. It allows you to work with various spatial data formats, such as shapefiles, GeoJSON, and WKT.
-   **tmap:** This package is designed for creating thematic maps in R. It offers a high-level interface for customizing maps with different styles, colors, and labels.
-   **tidyverse:** This is a collection of R packages that work together to provide a consistent and efficient data analysis workflow. It includes packages like dplyr for data manipulation, ggplot2 for data visualization, and readr for reading data files.  
-   **spatstat:** This package is specifically for spatial point pattern analysis. It provides tools for analyzing the distribution and patterns of points in a spatial region.
-   **raster:** This package is used for working with raster data, such as images and digital elevation models. It allows you to read, write, manipulate, and analyze raster data.
-   **sparr:** This package is for spatial point pattern analysis with regularly spaced data. It provides functions for estimating intensity functions, clustering, and other spatial point pattern analysis tasks.
-   **animation:** This package provides tools for creating animations in R. It can be used to visualize changes in spatial data over time or to create interactive visualizations.
-   **av:** This package is for working with AV files (audio-video files). It can be used to export animations or visualizations as videos.

```{r}
pacman::p_load(sf,tmap,tidyverse,spatstat,raster,sparr,animation,av,magick,htmltools)
```

# Importing data

-   **Conflict data:**

    -   `read.csv("data/acled_mm.csv")` reads a CSV file named "acled_mm.csv" and stores the data in a data frame named `conflict_data`. This file likely contains data on armed conflicts in Myanmar, such as the location, date, type, and actors involved in each conflict.

-   **Administrative boundary data:**

    -   `st_read(dsn="data", layer = "mmr_polbnda_adm1_250k_mimu_1")` reads a shapefile named "mmr_polbnda_adm1_250k_mimu_1" from the "data" directory and stores the data in a simple feature object named `boundary_data`. This shapefile likely contains the administrative boundaries of different regions or states in Myanmar, such as the borders of each state or province

```{r}
# Import the armed conflict data (2021-2024)
conflict_data <- read.csv("data/acled_mm.csv")

# Import administrative boundary data for Myanmar
boundary_data <- st_read(dsn="data",
                         layer = "mmr_polbnda_adm1_250k_mimu_1")

```

```{r}
summary(conflict_data)
```

The dataset appears to contain information on armed conflicts in Myanmar from 2021 to 2024. It includes various attributes that can be used for geospatial analysis:

**Location Data:**

-   **latitude and longitude:** These coordinates can be used to plot the locations of conflict events on a map.

-   **geo_precision:** This indicates the accuracy of the location data.

-   **admin1, admin2, admin3:** These likely represent administrative divisions such as states, districts, and townships, which can be used for spatial analysis at different scales.

**Event Data:**

-   **event_date, year, time_precision:** These attributes provide information about the date and time of the event.

-   **disorder_type, event_type, sub_event_type:** These categories describe the nature of the conflict event.

-   **actor1, actor2:** These likely refer to the actors involved in the conflict.

-   **interaction:** This might indicate the type of interaction between the actors (e.g., clashes, attacks).

-   **fatalities:** This provides information about the number of casualties.

```{r}
summary(boundary_data)
```

The output from `summary(boundary_data)` provides information about the structure and content of the administrative boundary data:

**Column Names:**

-   **OBJECTID:** This column likely represents a unique identifier for each administrative unit.

-   **ST:** This column might contain the name or abbreviation of the state or region.

-   **ST_PCODE:** This could be a numerical or alphanumeric code associated with the state.

-   **ST_RG:** This might refer to a specific region or district within the state.

-   **ST_MMR:** This could be another identifier or code related to the state.

-   **PCode_V:** This column might contain a postal code or other location identifier.

**Data Types:**

-   **OBJECTID:** Numeric

-   **ST, ST_PCODE, ST_RG, ST_MMR, PCode_V:** Character

**Summary Statistics:**

-   **Min., 1st Qu., Median, Mean, 3rd Qu., Max.:** These statistics provide information about the distribution of values in the `OBJECTID` and `PCode_V` columns.

## Cleaning data

```{r}
# Filter the dataset for the relevant event types
conflict_filtered <- conflict_data %>%
  filter(event_type %in% c("Battles", "Explosions/Remote violence", "Strategic developments", "Violence against civilians"))

# Convert the date column to Date format
conflict_filtered <- conflict_filtered %>%
  mutate(event_date = dmy(event_date))
```

Firstly we filter the `conflict_data` dataset to only include rows where the `event_type` is one of the specified types: "Battles", "Explosions/Remote violence", "Strategic developments", or "Violence against civilians". This is done to focus on the most relevant types of conflict events for the analysis.

Secondly, we convert the `event_date` column from its current format (presumably a character format) to a Date format using the `dmy()` function from the `lubridate` package. This is necessary for performing date-based calculations and visualizations.

```{r}
# Create a column for quarterly intervals (e.g., Q1 2021, Q2 2021, etc.)
conflict_filtered <- conflict_filtered %>%
  mutate(quarter = paste0(year(event_date), "-Q", quarter(event_date)))
```

Firstly,

-   `year(event_date)` extracts the year from the `event_date` column.

<!-- -->

-   `quarter(event_date)` extracts the quarter (1, 2, 3, or 4) from the `event_date` column

Next

`paste0(year(event_date), "-Q", quarter(event_date))` concatenates the year and quarter values with a hyphen and the letter "Q" in between. This creates a string representing the quarterly interval, such as "2021-Q1" or "2022-Q3".

Finally,we add the newly created `quarter` column to the `conflict_filtered` dataset.

```{r}
# Convert conflict data into an `sf` object with geometry
conflict_sf <- st_as_sf(conflict_filtered, coords = c("longitude", "latitude"), crs = 4326)

# Ensure the boundary data and conflict data have the same CRS (coordinate reference system)
conflict_sf <- st_transform(conflict_sf, st_crs(boundary_data))
```

We convert the `conflict_filtered` data frame into a simple feature object (`sf` object) using the `st_as_sf()` function from the `sf` package.

-   The `coords` argument specifies the columns containing the longitude and latitude coordinates (in this case, "longitude" and "latitude").

-   The `crs` argument sets the coordinate reference system (CRS) of the resulting `sf` object to 4326, which is the World Geodetic System 1984 (WGS84) coordinate system.

To ensure the matching CRS, we line transform the CRS of the `conflict_sf` object to match the CRS of the `boundary_data` object. This is important for spatial analysis and visualization, as the two datasets must have the same CRS to be combined or compared.

The `st_transform()` function is used to perform the CRS transformation. The `st_crs(boundary_data)` argument extracts the CRS of the `boundary_data` object.

# Visualizing and understanding Data

```{r}
# Plot spatial distribution of conflict events with boundary
ggplot() +
  geom_sf(data = boundary_data, fill = "white", color = "gray") +
  geom_sf(data = conflict_sf, aes(color = event_type), alpha = 0.6) +
  scale_color_manual(values = c("Battles" = "red", 
                                "Explosions/Remote violence" = "orange", 
                                "Strategic developments" = "blue", 
                                "Violence against civilians" = "purple")) +
  theme_minimal() +
  labs(title = "Spatial Distribution of Conflict Events (2021-2024)",
       color = "Event Type")
```

**Key Observations:**

1.  **Uneven Distribution:** The conflict events are not evenly distributed across Myanmar. There are regions with a higher concentration of events, particularly in the northern and eastern parts of the country.

2.  **Cluster Formation:** The map reveals clusters of conflict events, suggesting that certain areas have experienced more intense conflict activity.

3.  **Spatial Patterns:** The distribution of events appears to follow certain patterns, potentially influenced by geographical factors, historical grievances, or resource competition.

4.  **Event Type Variation:** The use of different colors to represent different event types allows for a visual comparison of their spatial distributions. This reveals that certain types of events may be more prevalent in specific regions.

**Analysis and Explanation:**

-   **Regional Hotspots:** The concentration of conflict events in the northern and eastern regions of Myanmar indicates that these areas have been particularly affected by the ongoing conflict. This could be due to a variety of factors, such as the presence of ethnic minorities seeking greater autonomy, competition for resources, or historical grievances.

-   **Spatial Patterns:** The clustering of events suggests that conflict activity may be concentrated in areas with specific characteristics, such as border regions, resource-rich areas, or areas with strategic importance.

-   **Event Type Distribution:** The visual representation of different event types provides insights into the nature of the conflict in different regions. For example, if battles are more prevalent in certain areas, it may suggest a more intense military confrontation.

-   **Potential Causal Factors:** The spatial distribution of conflict events can be analyzed in conjunction with other geospatial data, such as population density, land use, or proximity to natural resources, to identify potential causal factors.

**Overall, the map provides a valuable visual representation of the spatial dynamics of the conflict in Myanmar.** It highlights areas of intense conflict activity, reveals potential patterns in the distribution of events, and allows for a comparison of different event types. By analyzing these spatial patterns, researchers and policymakers can gain a better understanding of the underlying causes and consequences of the conflict, and develop more effective strategies for addressing it.

```{r}
# Temporal distribution of conflict events by quarter
ggplot(conflict_filtered, aes(x = quarter, fill = event_type)) +
  geom_bar(position = "stack") +
  scale_fill_manual(values = c("Battles" = "red", 
                               "Explosions/Remote violence" = "orange", 
                               "Strategic developments" = "blue", 
                               "Violence against civilians" = "purple")) +
  theme_minimal() +
  labs(title = "Temporal Distribution of Conflict Events by Quarter (2021-2024)",
       x = "Quarter", y = "Number of Events",
       fill = "Event Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Key Observations:**

1.  **Fluctuations:** The number of conflict events fluctuates over time, with periods of increased and decreased activity.

2.  **Seasonal Patterns:** Some event types may exhibit seasonal patterns, with higher or lower occurrences during specific quarters of the year.

3.  **Event Type Dominance:** Certain event types may be more prevalent than others during specific time periods.

**Analysis and Explanation:**

-   **Temporal Trends:** The chart reveals the overall trend of conflict activity over the four-year period. For example, there may be a general increase or decrease in the overall number of events.

-   **Seasonal Patterns:** Analyzing the fluctuations within each year can help identify any seasonal patterns. For instance, certain types of events may be more common during specific seasons due to factors such as weather conditions, agricultural activities, or religious holidays.

-   **Event Type Dynamics:** The chart allows for a comparison of the temporal trends of different event types. This can provide insights into the evolving nature of the conflict and the changing tactics of the involved parties.

-   **Potential Causal Factors:** The temporal trends can be analyzed in conjunction with other factors, such as political developments, economic conditions, or external influences, to identify potential causal factors.

**Overall, the chart provides a valuable visual representation of the temporal dynamics of the conflict in Myanmar.** It highlights periods of increased or decreased activity, reveals potential seasonal patterns, and allows for a comparison of different event types. By analyzing these temporal trends, researchers and policymakers can gain a better understanding of the factors driving the conflict and develop more effective strategies for addressing it.

```{r}
# Line plot for number of events over time
conflict_summary <- conflict_filtered %>%
  group_by(quarter, event_type) %>%
  summarize(n = n())

ggplot(conflict_summary, aes(x = quarter, y = n, color = event_type, group = event_type)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Battles" = "red", 
                                "Explosions/Remote violence" = "orange", 
                                "Strategic developments" = "blue", 
                                "Violence against civilians" = "purple")) +
  theme_minimal() +
  labs(title = "Trends in Conflict Events by Event Type (2021-2024)",
       x = "Quarter", y = "Number of Events",
       color = "Event Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Key Observations:**

1.  **Varied Trends:** Different event types exhibit distinct temporal patterns. Some show a steady increase, while others fluctuate or decline.

2.  **Peak Periods:** Certain quarters or years experienced peaks in specific event types, indicating periods of heightened conflict activity.

3.  **Interrelations:** The trends of different event types may be interrelated, suggesting that changes in one type of event may influence others.

**Analysis and Explanation:**

-   **Event Type Dynamics:** The chart reveals the unique temporal evolution of each event type. For example, "Battles" might have shown a significant increase in early 2022, followed by a decline, while "Violence against civilians" might have remained relatively consistent throughout the period.

-   **Peak Periods:** Identifying peak periods can help pinpoint critical junctures in the conflict. These periods might be associated with specific political events, military campaigns, or changes in the strategic landscape.

-   **Interrelations:** Analyzing the relationships between different event types can provide insights into the evolving nature of the conflict. For instance, an increase in "Strategic developments" might be followed by a subsequent rise in "Battles" as opposing forces adjust their strategies.

-   **Potential Causal Factors:** The temporal trends can be analyzed in conjunction with other factors, such as political developments, economic conditions, or external influences, to identify potential causal factors.

**Overall, the chart provides a valuable visual representation of the temporal dynamics of the conflict in Myanmar.** It highlights the unique trends of different event types, identifies peak periods, and suggests potential interrelations between events.

```{r}
conflict_by_region <- st_join(conflict_sf, boundary_data) %>%
  group_by(region = admin1) %>%  # Replace admin1Name with the correct region column
  summarize(n_conflicts = n())
# Histogram of conflict events by region
ggplot(conflict_by_region, aes(x = n_conflicts)) +
  geom_histogram(binwidth = 800, fill = "steelblue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Conflict Events by Administrative Region",
       x = "Number of Conflict Events",
       y = "Frequency (Number of Regions)",
       caption = "Each bin represents the number of conflict events in a region") +
  theme(plot.title = element_text(size = 16, face = "bold"))
```

**Key Observations:**

1.  **Skewed Distribution:** The distribution is skewed to the right, indicating that most regions have a relatively low number of conflict events. However, there are a few regions with significantly higher levels of conflict.

2.  **Mode:** The mode of the distribution appears to be between 3000 and 4000 conflict events, suggesting that this range is the most common among the regions.

3.  **Outliers:** There are a few outliers with a very high number of conflict events, indicating that these regions have experienced significantly more intense conflict than the majority.

**Analysis and Explanation:**

-   **Regional Disparities:** The skewed distribution highlights the uneven impact of the conflict across different regions. While most regions have experienced relatively low levels of conflict, a few have been disproportionately affected.

-   **Mode:** The mode indicates the most common level of conflict intensity. This information can be used to characterize the typical level of conflict in the regions.

-   **Outliers:** The outliers represent regions with exceptional levels of conflict. These areas may require special attention in terms of conflict resolution, humanitarian assistance, and development initiatives.

-   **Policy Implications:** The histogram can inform policy decisions by providing insights into the distribution of conflict intensity. This information can help prioritize areas for intervention, allocate resources effectively, and develop targeted strategies for conflict resolution

# Plot KDE of open street of Myanmar

```{r}
# Ensure the boundary data and conflict data have the same CRS (projected to UTM zone 47N)
boundary_data_projected <- st_transform(boundary_data, crs = 32647)
conflict_sf_projected <- st_transform(conflict_sf, crs = 32647)

# Convert the boundary data to an owin object (window) for spatstat
boundary_win <- as.owin(st_geometry(boundary_data_projected))
```

`st_geometry(boundary_data_projected)` extracts the geometry column from the `boundary_data_projected` object. This geometry column contains the spatial representation of the administrative boundaries.

`as.owin(st_geometry(boundary_data_projected))` converts the extracted geometry into an `owin` object using the `as.owin()` function from the `spatstat` package.

An `owin` object is a special data structure used by the `spatstat` package to represent the observation window or region within which point patterns are analyzed.

```{r}
# Convert conflict_sf_projected to ppp object
coords <- st_coordinates(conflict_sf_projected)
conflict_ppp <- ppp(x = coords[,1], y = coords[,2], window = boundary_win)

# Perform KDE for a specific quarter (e.g., Q1 2021)
kde_q3_2021 <- density.ppp(conflict_ppp, sigma = bw.diggle)

# Convert KDE result into a raster object
kde_raster <- raster(kde_q3_2021)
```

`coords <- st_coordinates(conflict_sf_projected)` extracts the coordinates of the conflict events from the `conflict_sf_projected` object. The resulting coordinates are stored in a matrix with two columns: one for longitude and one for latitude.

`conflict_ppp <- ppp(x = coords[,1], y = coords[,2], window = boundary_win)` creates a `ppp` object (point pattern) using the extracted coordinates and the `boundary_win` object as the observation window. A `ppp` object is a data structure used by the `spatstat` package to represent point patterns.

`kde_q3_2021 <- density.ppp(conflict_ppp, sigma = bw.diggle)` performs KDE on the `conflict_ppp` object using the `density.ppp()` function from the `spatstat` package. KDE estimates the intensity function of the point pattern, which represents the probability of finding a point at a given location. The `sigma` argument specifies the bandwidth of the kernel used for estimation.

`kde_raster <- raster(kde_q3_2021)` converts the KDE result (which is a `im` object) into a raster object using the `raster()` function from the `raster` package. This allows for easier visualization and integration with other raster data.

```{r}
# Assign projection to the raster
crs(kde_raster) <- "+proj=utm +zone=47 +datum=WGS84 +units=m +no_defs"

tm_shape(boundary_data_projected) + 
  tm_borders() + 
  tm_shape(kde_raster) +
  tm_raster(palette = "-Blues",style = "cont", alpha = 0.6) +
  tm_layout(title = "KDE of Armed Conflict Events (Q3 2021)")
```

## Understanding KDE

**1. Clustering:**

-   **Northern Myanmar:** The map shows a significant cluster of armed conflict events in the northern part of the country, particularly in Kachin and Shan states. This region has been a hotspot for conflict for many years due to the presence of various ethnic armed organizations.

-   **Western Myanmar:** Another cluster of events is observed in the western part of the country, along the border with Bangladesh. This area has been affected by conflict between the Myanmar government and the Rohingya Muslim minority.

**2. Dispersal:**

-   **Central Myanmar:** While there are fewer armed conflict events in central Myanmar compared to the north and west, there is still a noticeable dispersal of incidents throughout the region. This suggests that conflict is not confined to specific areas but can occur in various parts of the country.

**3. Proximity to Borders:**

-   **China and India:** The map highlights the potential influence of neighboring countries, particularly China and India, on the distribution of armed conflict events. The clusters in the north and west of Myanmar coincide with areas near the borders with these countries, suggesting that cross-border activities or support may be contributing to the conflict.

**4. Geographic Features:**

-   **Mountains and Rivers:** The spatial patterns may also be influenced by geographic features such as mountains and rivers. These can serve as natural barriers or routes for movement, affecting the accessibility of certain areas and the dynamics of conflict.

Overall, the KDE map provides valuable insights into the spatial distribution of armed conflict events in Myanmar during Q3 2021. It highlights areas of high intensity conflict, areas with a more dispersed occurrence of incidents, and the potential influence of geographic factors and neighboring countries.

# Plotting 2nd Order Spatial Point Pattern

```{r}
#| eval: false
# Take a random sample of 1000 events if the dataset is large
set.seed(123)  # Ensure reproducibility
sampled_conflict_ppp <- conflict_ppp[sample(1:npoints(conflict_ppp), 1000)]
# Compute Ripley's K function to analyze clustering or dispersion
K <- Kest(sampled_conflict_ppp, correction = "Ripley")

# Plot Ripley's K function
plot(K, main = "Ripley's K for Conflict Events (Q3 2021)")
```

## Understanding K-Test

Ripley's K function is a powerful tool in spatial statistics used to analyze point patterns and detect clustering or dispersion at various scales. Let's break down the key elements and implications of this graph

1.  **Clustering Evidence**: The K_iso(r) line consistently lies above the K_pois(r) line across all distances. This indicates significant clustering of conflict events at all spatial scales examined, suggesting that conflicts in Myanmar during this period were not randomly distributed but concentrated in specific areas.

2.  **Scale-Dependent Clustering**: The divergence between K_iso(r) and K_pois(r) increases with distance, implying that the strength of clustering becomes more pronounced at larger scales. This suggests the presence of both localized hotspots and broader regional patterns of conflict.

3.  **Non-Stationarity**: The continuous upward trend of both functions without leveling off indicates that the conflict process is not stationary. This implies that the intensity and distribution of conflicts vary across different regions of Myanmar.

4.  **Intensity of Clustering**: The steep curve of K_iso(r) compared to K_pois(r) suggests a high degree of clustering, indicating that some areas experienced significantly more conflict events than others.

5.  **Large-Scale Patterns**: The persistent clustering at large distances (up to 225,000 units) implies that the conflict has a broad spatial reach, potentially affecting large portions of the country or showing patterns of conflict that extend across significant geographic areas.

6.  **Implications for Conflict Dynamics**:

    -   The clustering pattern suggests that conflicts may be concentrated around specific strategic locations, ethnic territories, or resource-rich areas.

    -   The increasing divergence at larger scales could indicate a diffusion effect, where conflicts in one area may influence or spread to neighboring regions.

    -   The non-random distribution implies that there are underlying factors (e.g., political, ethnic, economic) driving the spatial patterns of conflict.

7.  **Methodological Considerations**: The use of K_iso(r) suggests that isotropy corrections have been applied, accounting for potential directional biases in the data and providing a more robust analysis of the spatial patterns.

    In conclusion, this Ripley's K function analysis reveals a highly clustered pattern of conflict events in Myanmar during Q3 2021, with clustering intensifying at larger spatial scales. This information is crucial for understanding the spatial dynamics of the conflict, which can inform targeted intervention strategies, resource allocation for peacekeeping efforts, and policy decisions aimed at conflict resolution and prevention. Further analysis combining this spatial information with temporal trends, conflict types, and socio-political factors would provide a more comprehensive understanding of the conflict situation in Myanmar.

# Plotting quarterly KDE layers

```{r}
# Unique quarters
quarters <- unique(conflict_filtered$quarter)

# Define and apply the projected CRS
utm_crs <- st_crs(32647)  # UTM Zone 48N
par(mfcol=c(2,3))
for (q in quarters) {
   # Filter data for the current quarter
  conflict_q <- conflict_filtered %>% 
    filter(quarter == q)
  # Add jitter to the longitude and latitude to avoid duplicated points
conflict_q <- conflict_q %>% 
  mutate(jittered_longitude = jitter(longitude, factor = 0.0001),
         jittered_latitude = jitter(latitude, factor = 0.0001))

# Convert to sf object using jittered coordinates
conflict_sf_q <- st_as_sf(conflict_q, coords = c("jittered_longitude", "jittered_latitude"), crs = 4326)
  
  # Transform to projected CRS
  conflict_sf_q_projected <- st_transform(conflict_sf_q, crs = utm_crs)
  
  # Convert boundary data to projected CRS and to owin object
  #boundary_data_projected <- st_transform(boundary_data, crs = utm_crs)
  #boundary_win <- as.owin(st_as_sf(boundary_data_projected))
  
  # Clip the jittered points to the boundary data
conflict_sf_q_projected <- st_intersection(conflict_sf_q_projected, boundary_data_projected)
  
  # Extract coordinates and convert to ppp object
  coords <- st_coordinates(conflict_sf_q_projected)
  conflict_ppp_q <- ppp(x = coords[,1], y = coords[,2], window = boundary_win)
  
  # Compute KDE
  kde <- density.ppp(conflict_ppp_q, sigma = bw.diggle)
  
  # Plot KDE with tmap
 plot(kde, main = paste("KDE for", q))
  
  
  
}
```

## Preparing data for quarterly KDE

-   **`quarters <- unique(conflict_filtered$quarter)`**

    -   This line extracts the unique values from the `quarter` column in the `conflict_filtered` dataset and stores them in the variable `quarters`.

    -   `unique()` is a function that returns only the distinct values from a given vector or column, so this line is essentially finding all the unique quarters present in your data.

-   **`utm_crs <- st_crs(32647)`**

    -   This line defines a coordinate reference system (CRS) using the `st_crs()` function from the `sf` package in R.

    -   `32647` refers to UTM Zone 48N, which is a specific coordinate system used for mapping and spatial analysis. The `st_crs()` function is used to assign this CRS to spatial objects in your analysis.

    -   `utm_crs` is being assigned this CRS so it can be applied to spatial data to ensure proper alignment and analysis.

-   **`par(mfcol=c(2,3))`**

    -   This line is setting up the plotting area for your graphics.

    -   `par(mfcol=c(2,3))` specifies that the plotting area should be divided into a 2x3 grid (2 rows and 3 columns) of panels for displaying multiple plots.

    -   This allows you to create multiple plots in one view, making it easier to compare different visualizations side-by-side.

<!-- -->

-   **`for (q in quarters)`**

    -   This line initiates a `for` loop that iterates over each unique value in the `quarters` vector.

    -   `q` is a loop variable that takes on the value of each unique quarter in turn.

-   **`conflict_q <- conflict_filtered %>% filter(quarter == q)`**

    -   Inside the loop, this line filters the `conflict_filtered` dataset to include only the rows where the `quarter` column matches the current value of `q`.

    -   `%>%` is the pipe operator from the `dplyr` package, which allows for chaining operations. In this case, it is used to pass the `conflict_filtered` dataset to the `filter()` function.

    -   `filter(quarter == q)` is a `dplyr` function that selects rows from `conflict_filtered` where the `quarter` column equals the current quarter value (`q`).

    -   `conflict_q` is a new variable that contains this filtered subset of the data for the current quarter.

**`mutate()` Function:**

-   **`jittered_longitude = jitter(longitude, factor = 0.0001)`**

    -   **`jitter(longitude, factor = 0.0001)`**: Applies the `jitter` function to the `longitude` column. The `factor = 0.0001` specifies a very small amount of noise to add, which spreads out points slightly.

    -   **`jittered_longitude`**: The new column name where the jittered values will be stored.

-   **`jittered_latitude = jitter(latitude, factor = 0.0001)`**

    -   **`jitter(latitude, factor = 0.0001)`**: Similar to the longitude, this adds jitter to the `latitude` column.

    -   **`jittered_latitude`**: The new column name where the jittered latitude values will be stored.

-   **Magnitude of Jitter:**

    -   The `factor` parameter in the `jitter()` function controls the amount of random noise added to your data. A smaller factor (like `0.0001`) results in minimal jitter, while a larger factor increases the amount of noise.

    -   For geographic coordinates, which are often measured in degrees, a very small factor is used to avoid significantly altering the positions while still making slight adjustments to avoid overlapping points.

-   **Scale of Data:**

    -   Geographic coordinates (latitude and longitude) are typically measured in decimal degrees. In this context, `0.0001` degrees is a small enough value to make minor adjustments without distorting the geographic locations significantly.

    -   For example, `0.0001` degrees of latitude or longitude is approximately equal to about 10 meters at the equator. This level of jitter is generally sufficient to visually separate closely spaced points without having a noticeable impact on their relative positions.

-   **Precision and Clarity:**

    -   When dealing with precise spatial data, such as GPS coordinates, it’s important to apply just enough jitter to improve visualization or analysis without introducing misleading inaccuracies.

    -   A factor of `0.0001` provides a balance between enough jitter to differentiate points and maintaining the precision of the original data

**`st_intersection()` Function:**

-   **Inputs:**

    -   `conflict_sf_q_projected`: The spatial points with jitter applied.

    -   `boundary_data_projected`: The spatial boundary or polygon that defines the area of interest.

-   **Operation:** `st_intersection()` calculates the geometric intersection between the points and the boundary. Points outside the boundary are discarded, and only those within the boundary are kept.

-   **Result:** `conflict_sf_q_projected` is updated to include only those points that lie within the boundary.

**Extracting Coordinates:**

-   **`st_coordinates()` Function:** This function extracts the coordinates from the `sf` object. It returns a matrix with columns for the x (longitude) and y (latitude) coordinates.

-   **Result:** `coords` is a matrix where `coords[,1]` contains the x-coordinates and `coords[,2]` contains the y-coordinates of the points within the boundary.

**Converting to `ppp` Object:**

-   **Arguments:**

    -   `x = coords[,1]`: The x-coordinates (longitudes) of the points.

    -   `y = coords[,2]`: The y-coordinates (latitudes) of the points.

    -   `window = boundary_win`: Defines the spatial window or boundary within which the points are analyzed. This should be a `psp` or `owin` object representing the area of interest.

-   **Result:** `conflict_ppp_q` is a `ppp` object that represents the spatial pattern of points within the defined boundary. This object is used for further spatial analysis and modeling.

## **Compute KDE**

**`density.ppp()` Function:**

-   **Purpose:** Computes the Kernel Density Estimate of point patterns. KDE is used to estimate the intensity (or density) of points in a spatial domain.

-   **Arguments:**

    -   `conflict_ppp_q`: The `ppp` object containing your spatial point data within the boundary.

    -   `sigma = bw.diggle`: Specifies the bandwidth for the KDE. `bw.diggle` is often a bandwidth parameter computed using methods such as Diggle’s method for optimal bandwidth selection. It controls the smoothness of the density estimate. A smaller bandwidth results in a more detailed but noisier density estimate, while a larger bandwidth provides a smoother, more generalized estimate.

-   **Result:** `kde` is an object representing the KDE of your point pattern, which can be visualized as a spatial density map.

## Understanding Quarterly KDE layers

The KDE maps provide insights into the spatial distribution of armed conflict events in Myanmar over different time periods:

**Overall Observations for 2021 Q3 - 2022 Q4:**

-   **Persistent Hotspots:** Certain regions in Myanmar consistently appear to have higher densities of armed conflict events throughout the time periods represented in the maps. These include the northern regions, particularly Kachin and Shan states, as well as areas near the western border with Bangladesh.

-   **Fluctuations:** While there are persistent hotspots, the intensity of conflict within these regions can fluctuate over time. Some maps show more concentrated clusters of events, while others indicate a more dispersed distribution.

**Specific Time Periods:**

-   **2021-Q3 and 2021-Q4:** These maps show a relatively high concentration of events in the northern part of Myanmar, particularly in Kachin and Shan states. There is also a noticeable cluster of events in the western region near the Bangladesh border.

-   **2022-Q1 and 2022-Q2:** The intensity of conflict seems to have decreased slightly in these quarters compared to the previous year. However, the northern and western regions remain hotspots, with a more dispersed distribution of events observed in some areas.

-   **2022-Q3 and 2022-Q4:** The maps for these quarters indicate a resurgence of conflict in the northern regions, with particularly high densities of events observed in Kachin state. The western region also continues to experience armed conflict, although the intensity may have fluctuated.

**Additional Observations:**

-   **Geographic Features:** The spatial patterns may be influenced by geographic features such as mountains and rivers, which can serve as natural barriers or routes for movement.

-   **Neighboring Countries:** The proximity of Myanmar to China and India may also play a role in the distribution of conflict, as cross-border activities or support could influence the dynamics in border regions.

**Key Observations:**

-   **Overall Density:** The overall density of the data points appears to be relatively low across all time periods, suggesting a sparse distribution.

-   **Clustering:** There are some indications of clustering in certain areas, particularly in the central and southern regions of the map. This suggests that data points are more concentrated in these areas compared to others.

-   **Temporal Changes:** Comparing the maps across different time periods, we can observe some changes in the spatial distribution. For example, the density in the central region seems to have increased slightly from 2023-Q1 to 2023-Q2, then decreased again in subsequent quarters.

-   **Regional Variations:** There are also regional variations in the density. The northern and eastern regions tend to have lower densities compared to the central and southern areas.

**Overall Observations for 2023 Q1-2024 Q2:**

**Overall Density:**

-   **KDE for 2024-Q2:** The highest density appears to be concentrated in the central and southern regions of the map, with a decreasing density towards the northern and eastern borders

-   **KDE for 2023-Q4:** Similar to the previous map, the highest density is observed in the central and southern areas, with a gradual decrease towards the northern and eastern regions.

-   **KDE for 2023-Q2:** The density is relatively high in the central region, with a more dispersed distribution compared to the previous two maps.

-   **KDE for 2024-Q1:** The highest density is concentrated in the central region, with a gradual decrease towards the northern and eastern borders.

-   **KDE for 2023-Q3:** The density is relatively high in the central region, with a more dispersed distribution compared to the previous two maps.

-   **KDE for 2023-Q1:** The highest density is observed in the central region, with a gradual decrease towards the northern and eastern borders.

**Temporal Changes:**

-   **Comparison of 2023-Q1 to 2023-Q4:** The overall density appears to have increased in the central region between these two quarters.

-   **Comparison of 2023-Q2 to 2023-Q3:** The distribution of density seems to have shifted slightly, with a potential increase in the southern region and a decrease in the central region.

-   **Comparison of 2024-Q1 to 2024-Q2:** The density appears to have increased in the central region between these two quarters.

*Overall, the KDE maps reveal a complex and evolving spatial pattern of armed conflict in Myanmar. While certain regions remain persistent hotspots, the intensity and distribution of conflict can fluctuate over time, reflecting the dynamic nature of the conflict landscape.*

# Plotting quarterly 2nd order Spatial Point Pattern

```{r}
par(mfcol=c(2,4))
for (q in quarters) {
  # Filter data for the current quarter
  conflict_q <- conflict_filtered %>% 
    filter(quarter == q)
  
  # Add jitter to the longitude and latitude to avoid duplicated points
  conflict_q <- conflict_q %>% 
    mutate(jittered_longitude = jitter(longitude, factor = 0.0001),
           jittered_latitude = jitter(latitude, factor = 0.0001))

  # Convert to sf object using jittered coordinates
  conflict_sf_q <- st_as_sf(conflict_q, coords = c("jittered_longitude", "jittered_latitude"), crs = 4326)
  
  # Transform to projected CRS
  conflict_sf_q_projected <- st_transform(conflict_sf_q, crs = utm_crs)
  
  # Convert boundary data to projected CRS and to owin object
  boundary_data_projected <- st_transform(boundary_data, crs = utm_crs)
  boundary_win <- as.owin(st_as_sf(boundary_data_projected))
  
  # Clip the jittered points to the boundary data
  conflict_sf_q_projected <- st_intersection(conflict_sf_q_projected, boundary_data_projected)
  
  # Extract coordinates and convert to ppp object
  coords <- st_coordinates(conflict_sf_q_projected)
  conflict_ppp_q <- ppp(x = coords[,1], y = coords[,2], window = boundary_win)
  
  # Compute K-function
  K_result <- Kest(conflict_ppp_q, correction = "border")
  
  # Plot the K-function
  plot(K_result, main = paste("K-function for", q), legend = FALSE)
  
  # Optionally, add theoretical vs. observed plot lines (e.g., K-Function minus CSR)
  plot(K_result, . - r ~ r, main = paste("K-test deviation for", q))
}
```

## Understanding quarterly 2nd order Spatial Point Pattern

**K-Function Plots:**

The K-function plots illustrate the average number of points within a distance r of a randomly chosen point in the study area. A K-function value greater than the expected value under complete spatial randomness (CSR) indicates clustering, while a value less than the expected value suggests dispersion.

-   **Q1:** The K-function curve remains consistently above the expected value at all distances, indicating a strong clustering pattern.

-   **Q2:** The K-function curve shows a similar pattern to Q1, with a strong clustering pattern observed at all distances.

-   **Q3:** The K-function curve lies above the expected value (dashed line) at small distances, suggesting clustering. However, at larger distances, the curve approaches the expected value, indicating a more random pattern.

-   **Q4:** Similar to Q3, the K-function curve initially shows clustering but then converges towards the expected value at larger distances.

**K-Test Deviation Plots:**

The K-test deviation plots provide a measure of the statistical significance of the observed clustering or dispersion. A deviation value significantly different from zero suggests a departure from CSR.

-   **Q1 and Q2:** The K-test deviation curves consistently show significant deviations from zero at all distances, confirming the strong clustering pattern observed in the K-function plots.

-   **Q3 and Q4:** The K-test deviation curves show some significant deviations from zero at smaller distances, indicating a significant clustering pattern. However, at larger distances, the deviations become less significant.

**Overall Conclusions:**

Based on the K-function and K-test deviation plots, the following conclusions can be drawn:

-   **Clustering:** The spatial point patterns for all four time periods exhibit a clustering pattern, with the strongest clustering observed in Q1 and Q2.

-   **Temporal Variation:** While clustering is evident in all periods, the intensity and extent of clustering may vary across time

-   **Statistical Significance:** The K-test deviation plots provide statistical evidence for the clustering patterns observed in the K-function plots.

These findings suggest that the spatial distribution of the events being analyzed is not random but rather exhibits a clustered pattern, which could be attributed to factors such as shared characteristics, common drivers, or spatial dependencies.

-   K-function Graphs:

    -   In all quarters, the observed K-function (solid black line) consistently lies above the theoretical K-function for complete spatial randomness (dashed red line).

    -   This indicates persistent clustering of conflict events across all time periods examine

-   K-test Deviation Graphs:

    -   These graphs show the difference between the observed and expected K-functions.

    -   The solid black line consistently lies above zero, further confirming significant clustering.

-   Temporal Trends:

    -   The overall shape and magnitude of clustering appear relatively consistent across quarters, suggesting a persistent pattern of conflict concentration.

    -   However, subtle variations in the steepness and height of the curves may indicate some fluctuations in the intensity of clustering over time.

-   Scale of Clustering:

    -   Clustering is evident at all scales examined (up to about 150,000 units, likely meters or kilometers).

    -   The increasing divergence between observed and expected K-functions at larger scales suggests that clustering becomes more pronounced over broader geographic areas.

-   Intensity of Conflict:

    -   The steepness of the observed K-function curves indicates high intensity of clustering, implying that conflicts are heavily concentrated in certain areas.

-   Spatial Non-stationarity:

    -   The continuous upward trend of the K-functions without leveling off suggests that the conflict process is not spatially stationary, meaning the intensity and distribution of conflicts vary across different regions of Myanmar.

-   Consistency Over Time:

    -   The similarity in patterns across quarters indicates that the overall spatial structure of the conflict has remained relatively stable from 2021 to 2023

    -   This suggests that the underlying factors driving conflict distribution (e.g., ethnic divisions, resource locations, strategic areas) have not significantly changed.

-   Implications for Conflict Analysis:

    -   The persistent clustering implies that conflict hotspots are relatively stable over time.

    -   This information could be crucial for targeting intervention efforts, allocating resources, and developing long-term conflict resolution strategies.

    -   The consistency of patterns might indicate entrenched conflict dynamics that require sustained, targeted approaches to address.

-   In conclusion, this comprehensive K-function analysis reveals a consistent pattern of significant spatial clustering in Myanmar's conflict events from 2021 to 2023. The persistence of these patterns over time suggests deeply rooted spatial dynamics in the conflict, likely tied to enduring geographic, political, or social factors. This information is invaluable for policymakers, humanitarian organizations, and peacekeeping efforts in developing strategies that address the specific spatial characteristics of the ongoing conflict in Myanmar.

# Preparing Data for Spatial Temporal KDE

```{r}
conflict_sf_spatemp<-conflict_sf_projected %>% 
  mutate(DayofYear = yday(event_date)) %>% 
  mutate(Month_num = month(event_date))
```

-   **`mutate(DayofYear = yday(event_date))`**:

    -   This creates a new column `DayofYear` that converts each event's `event_date` into a numeric value representing the day of the year (from 1 to 365 or 366 in a leap year).

    -   In spatial-temporal KDE, time is treated as a continuous dimension. Converting the event dates into day-of-year values allows you to model temporal variation across the calendar year in a continuous format. It helps identify patterns tied to specific times of the year, such as seasonal effects or yearly trends in conflict events.

-   **`mutate(Month_num = month(event_date))`**:

    -   This adds another column `Month_num` that extracts the month number from each event's `event_date` (values from 1 for January to 12 for December).

    -   Depending on the level of temporal granularity you wish to analyze, having the month number can be useful for grouping or summarizing events by months. While day-of-year is more granular, the month number allows for coarser temporal patterns, such as monthly trends or seasonal variations.

```{r}
conflict_sf_spatemp_month<-conflict_sf_spatemp %>% 
  dplyr::select(Month_num)
```

**`dplyr::select(Month_num)`**: This function selects only the `Month_num` column from the `conflict_sf_spatemp` dataset. The `dplyr::select()` function is used to subset data based on specific columns, and here you are choosing to isolate the `Month_num` column.

By selecting only `Month_num`, it can focus on analyzing temporal trends or summarizing events by month. This is useful when we want to focus on the temporal component without the spatial aspect.

```{r}
conflict_sf_spatemp_ppp <- as.ppp(conflict_sf_spatemp_month)
```

**`as.ppp()`**: This function is from the **spatstat** package, and it converts a dataset into a **planar point pattern (ppp)** object, which is a format used for spatial point pattern analysis. The `ppp` class contains spatial data (usually coordinates) and can be analyzed using methods for spatial point patterns (e.g., spatial density estimation, clustering).

```{r}
conflict_sf_spatemp_owin <-conflict_sf_spatemp_ppp[boundary_win]
```

**`conflict_sf_spatemp_owin`**: This new object will contain only the points from `conflict_sf_spatemp_ppp` that fall within the region defined by `boundary_win`.

```{r}
st_kde <- spattemp.density(conflict_sf_spatemp_owin)
```

The `spattemp.density()` function provides insights into how the density of conflict events changes across both space and time, giving it a better understanding of hotspots of activity over time.

# Plotting quarterly spatio-temporal KDE layers

```{r}
tims <- c(1,2,3,4,5,6,7,8,9,10,11,12)
par(mfcol=c(2,3))#2rows,3 col
for (i in tims){
 plot(st_kde,i,
       override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month",i))
} 
```

## Understanding quarterly spatio-temporal KDE layers

The provided maps represent Kernel Density Estimation (KDE) plots over a 12-month period. KDE is a statistical technique used to estimate the probability density function of a continuous random variable. In this context, it is likely being used to visualize the density or concentration of a particular phenomenon (e.g., population, events, or resources) across a geographic area over time.

-   Spatial pattern:

    -   The conflict appears to be primarily concentrated in the peripheral regions of Myanmar, particularly along the borders.

    -   The central plains area, which likely includes the capital region, shows consistently low conflict intensity throughout the period

-   Temporal evolution:

    -   Months 1-11 show relatively stable patterns of conflict distribution, with minor fluctuations in intensity.

    -   A noticeable change occurs in month 12, where there's a significant increase in conflict intensity, particularly in the northern and eastern border regions.

-   Hotspots:

    -   The northern region (likely Kachin State) consistently shows high conflict intensity throughout the period.

    -   The eastern border area (possibly Shan State) also maintains a persistent conflict presence.

    -   The western region (potentially Rakhine State) shows moderate conflict intensity that fluctuates over time.

-   Intensity changes:

    -   The color scale indicates that conflict intensity ranges from dark blue (low) to pink/red (high).

    -   Most months show localized areas of pink/red, indicating concentrated conflict zones.

    -   Month 12 exhibits a marked expansion of high-intensity areas, suggesting an escalation or spread of conflict.

-   Geographical factors:

    -   The conflict pattern aligns with Myanmar's ethnically diverse border regions, which have historically been prone to insurgencies and separatist movements.

    -   The relative peace in the central region may correspond to areas of stronger government control

# Perform 2nd-Order Spatio-temporal Point Patterns Analysis

```{r}
# Calculate spatio-temporal K-function
K_st <- Kest(conflict_sf_spatemp_ppp, correction = "border")

# Plot the spatio-temporal K-function
plot(K_st, main = "Spatio-temporal K-function")
```

**Key observations and interpretations**:

1.  **Conflict clustering**: The K_bord(r) function (solid black line) being consistently above the K_pois(r) function (dashed red line) indicates that conflict events in Myanmar are significantly clustered rather than randomly distributed. This suggests that violence is concentrated in certain areas rather than spread evenly across the country.

2.  **Scale of clustering**: The increasing divergence between K_bord(r) and K_pois(r) as distance increases implies that the clustering of conflict events becomes more pronounced at larger scales. This could indicate the presence of large regional conflict zones or that the conflict has a tendency to spread over time.

3.  **Persistent clustering**: The continuous upward trend of K_bord(r) without leveling off suggests that the clustering of conflict events persists even at large distances (up to 120,000 units, likely kilometers). This implies that the conflict in Myanmar has a broad spatial reach and may affect large portions of the country.

4.  **Intensity variation**: The steeper curve of K_bord(r) compared to K_pois(r) indicates that the intensity of conflict events varies significantly across different regions. Some areas likely experience much higher concentrations of conflict than others.

5.  **Conflict diffusion**: The increasing gap between K_bord(r) and K_pois(r) at larger distances might suggest a diffusion effect, where conflict in one area may influence or spread to neighboring regions over time.

6.  **Non-stationarity**: The lack of stabilization in the K-functions at larger distances indicates that the conflict process in Myanmar is not stationary. This suggests that the spatial patterns of conflict are dynamic and may be changing over time or varying across different regions of the country.

7.  **Complex spatial dynamics**: The shape of the K_bord(r) function, with its increasing slope, points to complex spatial dynamics in the conflict. This could reflect factors such as geography, ethnic distributions, resource locations, or strategic military considerations influencing where conflicts occur.

Implications for conflict analysis in Myanmar:

1.  **Hotspot identification**: The analysis clearly shows that conflict events are not random, allowing analysts to focus on identifying and studying specific conflict hotspots.

2.  **Conflict prediction**: The persistent clustering at various scales might be used to develop predictive models for potential conflict spread or escalation.

This spatio-temporal analysis provides a crucial quantitative foundation for understanding the complex dynamics of conflict in Myanmar, offering valuable insights for policymakers, humanitarian organizations, and peace negotiators working in the region.

# Preparing Data for Daily Spatial Temporal KDE

```{r}
conflict_sf_spatemp_day<-conflict_sf_spatemp %>% 
  dplyr::select(DayofYear)
```

**`dplyr::select(DayofYear)`**: This function selects only the DayofYearcolumn from the `conflict_sf_spatemp` dataset. The `dplyr::select()` function is used to subset data based on specific columns, and here you are choosing to isolate the DayofYearcolumn.

By selecting only DayofYear, it can focus on analyzing temporal trends or summarizing events by day of year. This is useful when we want to focus on the temporal component without the spatial aspect.

```{r}
conflict_sf_spatemp_day_ppp <- as.ppp(conflict_sf_spatemp_day)
```

```{r}
conflict_sf_spatemp_day_owin <-conflict_sf_spatemp_day_ppp[boundary_win]
```

```{r}
day_kde <- spattemp.density(conflict_sf_spatemp_day_owin)
```

# Plotting daily spatio-temporal KDE layers

```{r}
temp_dir <- "data/framesDay"

# Ensure the directory exists and create it if not
if (!dir.exists(temp_dir)) {
  dir.create(temp_dir, recursive = TRUE)
}

days <- 1:365
# Generate frames and save them as PNG images
for (i in days) {
  png_filename <- file.path(temp_dir, sprintf("frame_%03d.png", i))
  tryCatch({
    png(filename = png_filename)
    # Generate the plot for day i
    plot(day_kde, tselect = i, 
         override.par = FALSE,
         fix.range = TRUE,
         main = paste("KDE at day", i))
    title(xlab = "X", ylab = "Y")
  }, error = function(e) {
    message("Error generating frame ", i, ": ", e$message)
  }, finally = {
    dev.off()  # Ensure the device is closed regardless of success or failure
  })
}

frame_files <- list.files(temp_dir, pattern = "*.png", full.names = TRUE)

# Create the animation from the frames using magick
frame_images <- image_read(frame_files)
animation <- image_animate(frame_images, fps = 100)

# Save the animation as a GIF
gif_filename <- "spattemp_kde_day.gif"
image_write(animation, gif_filename)

```

```{r}
# Display the GIF in R's Viewer tab
browsable(tags$img(src = gif_filename))
```

## Understanding daily spatio-temporal KDE layers

-   Spatial **Distribution**: The image shows a KDE heat map overlaid on what appears to be the geographical outline of Myanmar. The intensity of the conflict or related events is represented by a color gradient, with yellows and reds indicating higher density/intensity, and blues showing lower density/intensity.

<!-- -->

-   **Focal** Point: There's a clear hotspot in the central region of the country, likely corresponding to a major city or strategic area. This suggests a concentration of conflict-related events or intensity in this specific location.

-   **Gradients**: The intensity radiates outward from the central hotspot, gradually diminishing towards the periphery. This pattern suggests a conflict with a clear epicenter but with ripple effects or related events spreading to surrounding areas.

<!-- -->

-   **Regional Variation**s: While the central area shows the highest intensity, there are smaller pockets of increased activity visible in other parts of the country, particularly in the north and southeast regions.

<!-- -->

-   **Scale and Interpretation**: The legend indicates values ranging from approximately 5e-15 to 3.5e-14. These small values likely represent probability densities, typical in KDE analysis, rather than absolute event counts.

# Conclusion & Reflection

### **1. Importance of Geospatial Analysis in Conflict Studies**

This exercise underscored how geospatial methods can reveal patterns that are often invisible in tabular data. The ability to visualize and analyze armed conflict events spatially and temporally offered a much clearer understanding of the scope and distribution of violence. From detecting the clustering of violence in key regions to observing temporal shifts in conflict zones, the KDE maps provided compelling evidence that allowed for deeper exploration of where and when conflict is concentrated. This kind of analysis can be instrumental in formulating responses, both in terms of immediate humanitarian aid and long-term peacebuilding.

### **2. The Challenges of Working with Conflict Data**

One of the reflections that emerged was the difficulty of working with conflict data, which often contains missing, duplicate, or imprecise points. In some cases, data points were rejected as they fell outside of defined boundaries, which reminded me of the importance of carefully pre-processing and cleaning the data before running the analysis. Additionally, geospatial data inherently comes with projection challenges, as the transition from geographic to projected coordinate systems is necessary for certain spatial analyses.

The warnings about duplicated points also illustrated the limitations of relying solely on point data to represent conflict events. Some areas are so saturated with violence that distinct events overlap, making it challenging to separate individual incidents without introducing statistical artifacts. This was a valuable reminder of the need for rigorous data wrangling and methodological precision when working with sensitive and complex datasets.

### **3. The Potential of Geospatial Tools in Policy and Humanitarian Efforts**

This assignment also highlighted the immense potential of geospatial tools in shaping policy and guiding humanitarian interventions. By identifying hotspots and shifts in conflict intensity over time, this type of analysis can help direct aid to the regions most affected by violence, and in real time, assist governments and international organizations in conflict mitigation efforts. For instance, timely monitoring of violence against civilians could allow for rapid responses to prevent further harm in emerging conflict zones. Moreover, identifying persistent hotspots provides crucial information for peace negotiations or ceasefire agreements that are geographically focused on the most affected areas.
