---
title: "In Class Exercise 5"
author: "edwin"
date: "September 16,2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## Loading the package

```{r}
pacman::p_load(sf, spdep, knitr, tmap, tidyverse,GWmodel)
```

## Preparing data

```{r}
#|eval: false
hunan<-st_read(dsn = 'data/geospatial',
               layer='Hunan')
```

```{r}
#|eval: false
hunan_2012<-read.csv('data/aspatial/Hunan_2012.csv')
```

```{r}
hunan_sf<-left_join(hunan,hunan_2012) %>% 
  select(1:3,7,15,16,31,32)
```

### Save Derived data

```{r}
#|eval: false
write_rds(hunan_sf,"data/rds/hunan_sf.rds")
```

### Read Derived data

```{r}
#|echo: false
hunan_sf<-read_rds("data/rds/hunan_sf.rds")
```

## Converting to SpatialPolygonDataFrame

```{r}
hunan_sp<-hunan_sf %>% 
  as_Spatial()
```

## Geographically Weighted Summary Statistics with adaptive bandwidth

### Determine adaptive bandwidth

::: panel-tabset
## CV (Cross validation)

```{r}
bw_AIC<- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "CV",
                adaptive = TRUE,
                kernel = "bisquare",
                longlat = T)
```

## AIC

```{r}
bw_AIC<- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "AIC",
                adaptive = TRUE,
                kernel = "bisquare",
                longlat = T)
```
:::

```{r}
bw_AIC
```

### Computing geographically weighted summary statistics

```{r}
gwstat<-gwss(data=hunan_sp,
             vars="GDPPC",
             bw=bw_AIC,
             kernel="bisquare",
             adaptive = TRUE,
             longlat=T)
```

### Preparing the output data

code below is used to extract SDF data table form gwss object output from gwss().It will be converted into data.frame by using as.data.frame()

```{r}
gwstat_df<-as.data.frame(gwstat$SDF)
```

Next,cbind() is used to append the newly derived data.frame onto hunan_sf sf data.frame

```{r}
hunan_gstat<-cbind(hunan_sf,gwstat_df)
```

### Visualising geographically weighted summary statistics

```{r}
tm_shape(hunan_gstat)+
  tm_fill("GDPPC_LM",
          n=5,
          style="quantile")+
  tm_borders(alpha=0.5)+
  tm_layout(main.title="Distibution of geoprahically weighted mean",
            main.title.position = "center",
            main.title.size = 1.0,
            legend.text.size = 0.7,
            legend.height = 1.50,
            legend.width = 1.50,
            frame=TRUE)
```
